---
title: "Reproducible Research: Course Project 1"
author: "Robert Osterried"
date: "March 11, 2016"
output: 
  html_document:
    keep_md: true
---


## Loading and Preprocessing the Data
First, let's load the data (CSV in the working directory) and take a look . . .
```{r, echo=TRUE}
data <- read.csv('activity.csv', na.strings = "NA")
names(data)
head(data)
dim(data)
```

## What is Mean Total Number of Steps Taken per Day?
Sum the steps for each day and make a histogram:
```{r, echo=TRUE}
totalSteps <- aggregate(data[, 1], list(Date=data$date), sum)
hist(totalSteps$x, main = "Total Number of Steps Taken Each Day", xlab = "Total Steps",
     col = "red")
```

Mean of total number of steps per day:
```{r, echo=TRUE}
mean(totalSteps$x, na.rm = TRUE)
```

Median of total number of steps per day:
```{r, echo=TRUE}
median(totalSteps$x, na.rm = TRUE)
```

## What is the Average Daily Activity Pattern?

Average the steps for each five-minute interval across all days and make a time series plot:
```{r, echo=TRUE}
intervalSteps <- aggregate(data[, 1], list(Interval=data$interval), mean, na.rm=TRUE)
library(lattice)
xyplot(x ~ Interval, intervalSteps, type = "l", xlab = "5-Min Interval", 
       ylab = "Avg. Steps")
```

Determine which 5-minute interval, on average across all days, has the maximum number of steps:
```{r, echo=TRUE}
intervalSteps[intervalSteps$x == max(intervalSteps$x),]$Interval
```

## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?